# Reproducible Research: Peer Assessment 1
**Completed May, 18 2014**


## Loading and preprocessing the data
### 1) Load the data
```{r echo=TRUE}
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip")
unzip("activity.zip")
data_set<-read.csv("activity.csv", as.is = TRUE)
```
### 2) Process/transform the data (if necessary) into a format suitable for your analysis

```{r echo=TRUE}
#Coerce the date column to a Date data type.
data_set$date<-as.Date(data_set$date)

#Compute the actual minutes from the start of each day form the interval data and add it to the data set.
minutes<-((data_set$interval-data_set$interval%%100)/100*60)+data_set$interval%%100
data_set<-cbind(data_set,minutes)
```
___
## What is mean total number of steps taken per day?
### 1) Make a histogram of the total number of steps taken each day
```{r echo=TRUE}
#Create aggregate data frame. Omit NAs. (This is done by default but the parameter was included here for clarity)
total_steps_by_day<-aggregate(steps ~ date, data_set, sum, na.action = na.omit)
#Produce histogram
hist(total_steps_by_day$steps,main="Total Steps per Day",xlab = "Number of Steps", col = "LightSteelBlue")
```
### 2) Calculate and report the mean and median total number of steps taken per day
```{r echo=TRUE}
#Please note that NAs were omitted in the creation of the total_steps_by_day data frame.
mean_steps <- mean(total_steps_by_day$steps)
median_steps <- median(total_steps_by_day$steps)
```
**Results:**  
Mean Steps per Day: `r as.integer(mean_steps)`  
Median Steps per Day: `r as.integer(median_steps)`
___

## What is the average daily activity pattern?
### 1) Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo=TRUE}
#Create aggregate data frame.
average_steps_by_interval<-aggregate(steps ~ interval, data_set, mean, na.action = na.omit)
#produce plot
plot(average_steps_by_interval, type = "l", main = "Average Steps by Interval")
```

###2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
max_steps <- subset(average_steps_by_interval,average_steps_by_interval$steps == max(average_steps_by_interval$steps), select = interval)
```
**Results:**  
Interval with the most steps: `r as.integer(max_steps$interval)`

___
## Imputing missing values
### 1)Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r echo=TRUE}
missing_values<-sum(is.na((subset(data_set,is.na(steps),select = steps))))
```
**Results:**  
Number of missing values: `r missing_values`

###2) Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r echo=TRUE}
merged_data_set<-merge(data_set, average_steps_by_interval, by="interval")
merged_data_set1<-subset(merged_data_set,is.na(steps.x),select = c(steps.y, date, interval))
merged_data_set2<-subset(merged_data_set,!is.na(steps.x),select = c(steps.x, date, interval))
names(merged_data_set1)<-c("steps", "date", "interval")
names(merged_data_set2)<-c("steps", "date", "interval")
merged_data_set<-rbind(merged_data_set1,merged_data_set2)
merged_data_set<-merged_data_set[with(merged_data_set, order(date, interval)), ]
```


**Ran out of time here sorry**
___
## Are there differences in activity patterns between weekdays and weekends?
