# Reproducible Research: Peer Assessment 1
**Completed June, 15 2014**


## Loading and preprocessing the data
### 1) Load the data
```{r echo=TRUE}
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip")
unzip("activity.zip")
data_set<-read.csv("activity.csv", as.is = TRUE)
```
### 2) Process/transform the data (if necessary) into a format suitable for your analysis

```{r echo=TRUE}
#Coerce the date column to a Date data type.
data_set$date<-as.Date(data_set$date)

#Compute the actual minutes from the start of each day form the interval data and add it to the data set.
minutes<-((data_set$interval-data_set$interval%%100)/100*60)+data_set$interval%%100
data_set<-cbind(data_set,minutes)
```
___
## What is mean total number of steps taken per day?
### 1) Make a histogram of the total number of steps taken each day
```{r echo=TRUE}
#Create aggregate data frame. Omit NAs. (This is done by default but the parameter was included here for clarity)
total_steps_by_day<-aggregate(steps ~ date, data_set, sum, na.action = na.omit)
#Produce histogram
hist(total_steps_by_day$steps,main="Total Steps per Day",xlab = "Number of Steps", col = "LightSteelBlue")
```
### 2) Calculate and report the mean and median total number of steps taken per day
```{r echo=TRUE}
#Please note that NAs were omitted in the creation of the total_steps_by_day data frame.
mean_steps <- mean(total_steps_by_day$steps)
median_steps <- median(total_steps_by_day$steps)
sd_data<-sd(total_steps_by_day$steps)
```
**Results:**  
Mean Steps per Day: `r as.integer(mean_steps)`  
Median Steps per Day: `r as.integer(median_steps)`  
Standard Deviation: `r as.integer(sd_data)`  
___

## What is the average daily activity pattern?
### 1) Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo=TRUE}
#Create aggregate data frame.
average_steps_by_interval<-aggregate(steps ~ interval, data_set, mean, na.action = na.omit)
#produce plot
plot(average_steps_by_interval, type = "l", main = "Average Steps by Interval")
```

###2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
max_steps <- subset(average_steps_by_interval,average_steps_by_interval$steps == max(average_steps_by_interval$steps), select = interval)
```
**Results:**  
Interval with the most steps: `r as.integer(max_steps$interval)`

___
## Imputing missing values
### 1)Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r echo=TRUE}
missing_values<-sum(is.na((subset(data_set,is.na(steps),select = steps))))
```
**Results:**  
Number of missing values: `r missing_values`

### 2) Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r echo=TRUE}
#We will use the average steps to fill in missing values
merged_data_set<-merge(data_set, average_steps_by_interval, by="interval")
merged_data_set1<-subset(merged_data_set,is.na(steps.x),select = c(steps.y, date, interval))
merged_data_set2<-subset(merged_data_set,!is.na(steps.x),select = c(steps.x, date, interval))
names(merged_data_set1)<-c("steps", "date", "interval")
names(merged_data_set2)<-c("steps", "date", "interval")
merged_data_set<-rbind(merged_data_set1,merged_data_set2)
merged_data_set<-merged_data_set[with(merged_data_set, order(date, interval)),]
```

### 3)Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r echo=TRUE}
#See code above for actual table creation
missing_values<-sum(is.na((subset(merged_data_set,is.na(steps),select = steps))))
```
**Results:**  
Number of missing values: `r missing_values`


### 4) Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r echo=TRUE}
#Create aggregate data frame. 
total_steps_by_day<-aggregate(steps ~ date, merged_data_set, sum)
#Produce histogram
hist(total_steps_by_day$steps,main="Total Steps per Day",xlab = "Number of Steps", col = "LightSteelBlue")

mean_merged_steps <- mean(total_steps_by_day$steps)
median_merged_steps <- median(total_steps_by_day$steps)
sd_merged_data<-sd(total_steps_by_day$steps)

```
**Results:**  
Mean Steps per Day: `r as.integer(mean_merged_steps)`  
Median Steps per Day: `r as.integer(median_merged_steps)`  
Standard Deviation: `r as.integer(sd_merged_data)`  
There is no significant difference from the mean estimates from the first part of the assignment. This is due to using mean values to fill in the missing data.
The observed effects of using that data to fill in the missing values are a lower standard deviation and that the median has moved closer to the mean.


## Are there differences in activity patterns between weekdays and weekends?  

### 1) Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r echo=TRUE}
library(plyr)
merged_data_set$day_type <- factor(weekdays(as.Date(merged_data_set$date)) == "Sunday" | weekdays(as.Date(merged_data_set$date)) == "Saturday", labels = c("Weekday", "Weekend"))
total_steps_by_day_type <- ddply(merged_data_set, .(day_type, interval), summarize, total_steps = sum(steps))
```

### 2) Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r echo=TRUE}
library(lattice)
xyplot(total_steps ~ interval | day_type, data = total_steps_by_day_type, layout = c(1, 2), type = "l")
```

